---
title: 'R markdown: A tutorial'
author: "Paul Schneider & Sarah Bates"
date: "18/01/2022"
output: 'html_document'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Excercise

1. Change the format to html
2. Use subsetting to present the results from one country only
3. Include the table in the report. See if you can present the columns year, lifeExp, pop and gdpPercap only

# Dataframe

```{r, echo = F}

df_gapminder <- read.csv(file = "https://raw.githubusercontent.com/ScHARR-PHEDS/R4ScHARR/master/data/df_gapminder.csv")

#Subset to United Kingdom only
df_UK <- df_gapminder[df_gapminder$country == "United Kingdom",]

#subset to columns of interest
df_UK<-df_UK[,c("year", "lifeExp","pop", "gdpPercap")]  

# remove rownames
rownames(df_UK) <- c()

df_UK
```


## Including Plots

4. Add a plot with population over time in that country
5. Change the code so that the code chunk can't be seen in the report

```{r, echo=F, message=FALSE, warning = FALSE}


# Plot these on a line graph
library(ggplot2)

# line graph of population against year
ggplot(data=df_UK, aes(x=year, y=pop)) +
  geom_line()+
  geom_point()


```

##

5. Report the change in population between the year 1957 and 2007

The population has increased by `r toString(df_UK[12, "pop"] - df_UK[2,"pop"])` between the years 1957 and 2007.

The population has increased by `r toString(df_UK$pop[df_UK$year==2007] - df_UK$pop[df_UK$year==1957])` between the years 1957 and 2007.




