---
title: 'R markdown: A tutorial'
author: "Paul Schneider & Sarah Bates"
date: "18/01/2022"
output: 'html_document'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Excercise


# Average Life expectancy

```{r, echo = F}

df_gapminder <- read.csv(file = "https://raw.githubusercontent.com/ScHARR-PHEDS/R4ScHARR/master/data/df_gapminder.csv")

# 1. Subset the data frame to a single year

df_2007 <- df_gapminder[df_gapminder$year == 2007,]

# 2. Use a for loop to identify the mean life expectancy for each continent

# The function unique can be used to identify the unique value in a column
continents <- c(unique(df_2007$continent))

for (i in continents){
  
  df_x <- df_2007[df_2007$continent  == i, ]

  df_mean <- data.frame(i, mean(df_x$lifeExp)) 
  
  names(df_mean)<- NULL
  
  print(df_mean)
                                             
}

```


## Average GDP

```{r, echo=F, message=FALSE, warning = FALSE}


# 3. Create a custom function that present the mean GDP per capita for a certain continent and year.

f_gdpPercap <- function(continent, year){
  
  df_year <- df_gapminder[df_gapminder$year == year,]
  
  df_continent <- df_year[df_year$continent  == continent, ]
  
  df_mean <- data.frame(continent, Mean_GDP = (mean(df_continent$gdpPercap)))

  print(df_mean)
    
}

# 4. Use this function to print the mean GDP for each continent in year you previously selected

f_gdpPercap("Asia", 2007)
f_gdpPercap("Europe", 2007)
f_gdpPercap("Africa", 2007)
f_gdpPercap("Americas", 2007)
f_gdpPercap("Oceania", 2007)


```

## Maximum values

5. Add text that identifies the highest life expectancy and highest GDP per capita in the year you selected

The highest life expectancy was `r toString(max(df_2007$lifeExp))` in `r toString(df_2007$country[df_2007$lifeExp==max(df_2007$lifeExp)])`.

The highest GDP per capita was `r toString(max(df_2007$gdpPercap))` in `r toString(df_2007$country[df_2007$gdpPercap==max(df_2007$gdpPercap)])`.




